---
title: "Overinformativeness BDA"
output: html_notebook
---

# Set up environment

```{r}
library(tidyverse)
library(coda)
library(lme4)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10) 

source("./helper_scripts/model_helpers.R")
```


### Load in model results

```{r}
params <- read_csv("../models/typicality-params.csv")
params %>% group_by(parameter, value) %>% summarize(prob = sum(MCMCprob))
```

look at MAP, credible intervals...

```{r}
params %>% 
  group_by(parameter) %>%
  summarize(mode = estimate_mode(value),
            md_lo = round(HPDlo(value), 3),
            md_hi = round(HPDhi(value), 3))
```

```{r}
numSamples = nrow(params)/length(levels(params$parameter))
params.samples <- params[rep(row.names(params), exp(params$MCMCprob)*numSamples), 1:2]

#pdf(file="../writing/2016/journal-manuscript/figures/AnswererParamPosteriors.pdf",
        #width = 6, height = 3)
answererParamPosteriors = ggplot(params.samples, aes(x=value))+
  #geom_histogram(aes(y=..density..)) +
  
    geom_histogram(aes(y=..density..),
                   data =subset(params.samples, parameter == "alpha" ),
                   binwidth = .5, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter == "typWeight"),
                   binwidth = .25, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "cost_color" ),
                 binwidth = .5, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                 data=subset(params.samples, parameter == "cost_type"),
                 binwidth = .25, colour="black", fill="white") +

    geom_density(aes(y=..density..),
                 data =subset(params.samples, parameter == "alpha" ),
                 adjust = 1, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "typWeight"),
                 adjust = 5, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "cost_color"),
                 adjust = 5, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "cost_type"),
                 adjust = 5, alpha=.2, fill="#FF6666")+

    ggtitle("Parameter Posteriors (500 iterations)") +
    facet_grid(~parameter, scales = 'free') +
    theme_bw()
answererParamPosteriors
```

Looks like there's a super narrow range where the pragmatic model makes much much better predictions... 

### Predictives

# Import empirical data

```{r}
empirical = read.csv("../models/bdaInput/typicality/bda_data-raw.csv") %>%
  mutate(target_item = paste(t_color, t_item, sep = "_")) %>%
  mutate(color = ifelse(as.character(response) == as.character(t_color),1,0)) %>%
  mutate(item = ifelse(as.character(response) == as.character(t_item),1,0)) %>%
  mutate(color_item = ifelse(as.character(response) == as.character(target_item),1,0))

agr = empirical %>%
  select(color,item,color_item,condition,target_item) %>%
  gather(utterance,mentioned,-condition,-target_item) %>%
  group_by(condition,utterance,target_item) %>%
  summarise(Probability=mean(mentioned),
            cilow=ci.low(mentioned),
            cihigh=ci.high(mentioned)) %>%
  ungroup() %>%
  mutate(YMax = Probability + cihigh) %>%
  mutate(YMin = Probability - cilow) %>%
  select(condition, target_item, utterance, Probability, YMax, YMin)
agr = as.data.frame(agr)
```

# Import predictives data

```{r}
colors <- c("yellow","orange","red","pink", "green","blue","brown","black")
types <- c('apple', 'avocado', 'banana', 'carrot', 'pear', 'tomato', 'pepper')
predictive<-read.csv("../models/typicality-predictives.csv") %>%
  rowwise() %>%
  mutate(target_item = paste0(c(as.character(t_color), as.character(t_item)), collapse = '_')) %>%
  filter(response %in% c(as.character(t_color), as.character(t_item), target_item)) %>%
  mutate(utterance = ifelse(as.character(response) %in% colors, 'color', 
                            ifelse(as.character(response) %in% types, 'item',
                                   'color_item'))) %>%
  group_by(condition, utterance, target_item) %>%
  summarize(modelProbability=mean(logModelProb)) 
```

# Plot predictives w/ empirical data

```{r}
combined <- left_join(agr, predictive)
answer_plots = (ggplot(combined, aes(x = modelProbability, y = Probability))
  + geom_errorbar(aes(ymax = YMax, ymin = YMin), alpha = .2) 
  + geom_point(aes(colour = condition))
  + geom_abline(intercept = 0, slope = 1, linetype = "dotted")
  + scale_x_continuous(lim = c(0,1), breaks=c(0,.5,1))
  + ggtitle("Answerer Posterior Predictive")
  #+ geom_smooth(method = "lm")
  + xlab("Model predicted probability")
  + ylim(0,1)
  + facet_wrap(~ utterance)
  #+ geom_text(aes(label=utterance), size = 2)
  + ylab("Empirical Probability")
  + theme_bw()   
  + theme(text = element_text(size = 10),  axis.text.x = element_text(angle=90, vjust=1), aspect.ratio = 1)
)
answer_plots
#dev.off()
```

Look at R^2

```{r}
summary(lm(Probability ~ modelProbability, data = combined))
```

Plot qualitative pattern

```{r}
typ = read.csv(file="~/cogsci/projects/stanford/projects/overinformativeness/experiments/elisa_paper_relevant/norming_comp_object/results/data/meantypicalities.csv")
typ = typ[as.character(typ$Item) == as.character(typ$utterance),]
row.names(typ) = paste(typ$Color,typ$Item,sep="_")
combined$NounTypicality = typ[as.character(combined$target_item),]$MeanTypicality

ggplot(combined, aes(x=NounTypicality,y=modelProbability,color=Utterance)) +
  geom_point() +
  geom_smooth(method="lm") +
  facet_wrap(~conditionName)
```


Compare predictives with & without emperical lengths/weights

```{r}
predictiveComparison <-read_csv("./MAPneglCostPredictives.csv") %>%
  rename(empCostProb = prob) %>%
  left_join(read_csv("./MAPfixedcostPredictives.csv")) %>%
  rename(fixedCostProb = prob) %>%
  rowwise() %>%
  mutate(target_item = paste0(c(TargetColor, TargetType), collapse = '_')) %>%
  filter(value %in% c(TargetColor, TargetType, target_item)) %>%
  mutate(Utterance = ifelse(value %in% colors, 'color', 
                            ifelse(value %in% types, 'type',
                                   'color_type'))) %>%
  rename(conditionName = condition) %>%
  group_by(conditionName, Utterance, target_item) %>%
  summarize(empProb=mean(empCostProb), nonEmpProb = mean(fixedCostProb)) 

ggplot(predictiveComparison, aes(x = empProb, y = nonEmpProb)) +
  geom_point() +
  geom_text(aes(label=Utterance), size = 2)
```