---
title: "Overinformativeness BDA"
output: html_notebook
---

# Set up environment

```{r}
library(tidyverse)
library(coda)
library(lme4)
library(magrittr)
library(gridExtra)
library(ggnewscale)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10) 

# set working directory to directory of script
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)

source("./helper_scripts/model_helpers.R")

# Get color-blind friendly palette that also looks good in black and white
# http://dr-k-lo.blogspot.com/2013/07/a-color-blind-friendly-palette-for-r.html
cbPalette <- c("#000000", "#009E73", "#e79f00", "#9ad0f3", "#0072B2", "#D55E00", "#CC79A7", "#F0E442")
```

# Analyze color-size results

## Parameters

Read parameter data from BDA and compute MAPs and HDIs.
```{r}
params <- read_csv("../models/bdaOutput/colorSize_params.csv") %>%
  mutate(costDiff = colorCost - sizeCost) %>%
  mutate(typDiff = colorTyp - sizeTyp) %>%
  gather(parameter, value, alpha:typWeight, costDiff, typDiff)
  # gather(parameter, value, infWeight:typWeight, costDiff, typDiff)

params %>% 
  group_by(parameter) %>%
  summarise(mode = estimate_mode(value),
            md_lo = round(HPDlo(value), 3),
            md_hi = round(HPDhi(value), 3))
```

```{r}
params.samples <- getSamples(params)

labels <- c(colorTyp = "color", sizeTyp = "size")
typParamPosteriors = ggplot(params.samples %>% filter(parameter %in% c('colorTyp', 'sizeTyp')), aes(x=value))+
    geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter == "colorTyp"),
                   binwidth = .005, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "sizeTyp" ),
                 binwidth = .005, colour="black", fill="white") +

    geom_density(aes(y=..density..), alpha=.3,
                 data=subset(params.samples, parameter == "colorTyp"),
                 adjust = 2, infWeight=.2, fill="#FF6666")+
    geom_density(aes(y=..density..), alpha=.3,
                 data=subset(params.samples, parameter == "sizeTyp"),
                 adjust = 2, infWeight=.2, fill="#FF6666")+
    # ggtitle("Param Posteriors (2000 samples; lag = 10; burnin=10000)") +
    facet_grid(parameter ~. , scales = 'free', labeller=labeller(parameter = labels)) +
    # theme_bw() +
    xlim(.7, 1) +
    ylab("Density") +
    xlab("Semantic value") +
    theme(panel.border = element_rect(size=.2),
        plot.margin = unit(x = c(0.05, 0.02, 0.05, 0.05), units = "in"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .5)),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=4,r=4,b=4,l=4,unit="pt")))
typParamPosteriors

```

Clearly see that color semantic vaue is much greater than size (basically non-overlapping posteriors).


```{r}
labels <- c(colorCost = "color", sizeCost = "size")
costParamPosteriors = ggplot(params.samples %>% filter(parameter %in% c('colorCost', 'sizeCost')), aes(x=value))+
    geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter == "colorCost"),
                   binwidth = .02, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "sizeCost" ),
                 binwidth = .02, colour="black", fill="white") +

    geom_density(aes(y=..density..), alpha=.5,
                 data=subset(params.samples, parameter == "colorCost"),
                 adjust = 1, infWeight=.2, fill="#FF6666")+
    geom_density(aes(y=..density..), alpha=.5,
                 data=subset(params.samples, parameter == "sizeCost"),
                 adjust = 1, infWeight=.2, fill="#FF6666")+
    ylab("Density") +
  xlab("Cost") +

    # ggtitle("Param Posteriors (2000 samples; lag = 10; burnin=10000)") +
    facet_grid(parameter ~. , scales = 'free', labeller=labeller(parameter = labels)) +
    # theme_bw() +
    theme(panel.border = element_rect(size=.2),
        plot.margin = unit(x = c(0.05, 0.02, 0.05, 0.05), units = "in"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .5)),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=4,r=4,b=4,l=4,unit="pt")))
    #xlim(.5, 1)
costParamPosteriors
```

Stick together semantic value and cost plots for paper.
```{r}
# pdf(file="../writing/pics/exp1-paramposteriors.pdf",height=3.5,width=8)
# grid.arrange(typParamPosteriors,costParamPosteriors,nrow=1)
# dev.off()

tiff(file="../writing/pics/fig10.tiff",units="in",width=5.5,height=2.7,res=600)
grid.arrange(typParamPosteriors,costParamPosteriors,nrow=1)
dev.off()
```

Inferred difference in color and size semantic value always above zero:
```{r}
answererParamPosteriors = ggplot(params.samples %>% filter(parameter %in% c('typDiff')), aes(x=value))+
    geom_histogram(data=subset(params.samples, parameter == "typDiff"),
                   binwidth = .003, colour="black", fill="white") +

    ggtitle("Param Posteriors (2000 samples; lag = 10; burnin=10000)") +
    facet_grid(parameter ~. , scales = 'free') +
    theme_bw() 
    #xlim(.5, 1)
answererParamPosteriors
```

Inferred difference in color and size cost around zero:
```{r}
answererParamPosteriors = ggplot(params.samples %>% filter(parameter %in% c('costDiff')), aes(x=value))+
    geom_histogram(data=subset(params.samples, parameter == "costDiff"),
                   binwidth = .05, colour="black", fill="white") +

    ggtitle("Param Posteriors (2000 samples; lag = 10; burnin=10000)") +
    facet_grid(parameter ~. , scales = 'free') +
    theme_bw() 
    #xlim(.5, 1)
answererParamPosteriors
```

## Predictives

Load empirical data and get into right format.

```{r}
emp = read.csv("../models/bdaInput/colorSize/bda_data.csv")
colnames(emp) = c("ID","Trial","Condition","Size","Color","OtherColor","Item","Utterance")
empirical = emp %>%
  mutate(CondLength = nchar(as.character(Condition))) %>%
  mutate(SufficientDimension = 
           case_when( CondLength == 7 ~ "color",
                      CondLength == 6 ~ "size",
                      TRUE ~ "error"),
         NumDistractors = as.numeric(as.character(str_sub(Condition,CondLength-1,CondLength-1))),
         NumSameDistractors = as.numeric(as.character(str_sub(Condition,CondLength,CondLength))),
         NumDiffDistractors = NumDistractors-NumSameDistractors,
         Variation = NumDiffDistractors/NumDistractors) %>%
  select(SufficientDimension,NumDistractors,Variation,Utterance) %>%
  mutate(Color = Utterance == "color",
         Size = Utterance == "size",
         Size_Color = Utterance == "size_color") %>%
  group_by(SufficientDimension,NumDistractors,Variation) %>%
  summarise(ProbColor = mean(Color),ProbSize = mean(Size),ProbSize_Color = mean(Size_Color),CILowC=ci.low(Color),CIHighC=ci.high(Color),CILowS=ci.low(Size),CIHighS=ci.high(Size),CILowSC=ci.low(Size_Color),CIHighSC=ci.high(Size_Color)) %>%
  ungroup() %>%
  mutate(YMinSize=ProbSize-CILowS,YMaxSize=ProbSize+CIHighS,YMinColor=ProbColor-CILowC,YMaxColor=ProbColor+CIHighC,YMinSize_Color=ProbSize_Color-CILowSC,YMaxSize_Color=ProbSize_Color+CIHighSC)

tmp_prob = empirical %>%
  select(SufficientDimension,NumDistractors,Variation,ProbColor,ProbSize,ProbSize_Color) %>%
  gather(Utterance, Probability,-SufficientDimension,-NumDistractors,-Variation) %>%
  mutate(Utterance=gsub("Prob","",as.character(Utterance)))

tmp_ymin = empirical %>%
  select(SufficientDimension,NumDistractors,Variation,YMinColor,YMinSize,YMinSize_Color) %>%
  gather(Utterance, YMin,-SufficientDimension,-NumDistractors,-Variation) %>%
  mutate(Utterance=gsub("YMin","",as.character(Utterance)))

tmp_ymax = empirical %>%
  select(SufficientDimension,NumDistractors,Variation,YMaxColor,YMaxSize,YMaxSize_Color) %>%
  gather(Utterance, YMax,-SufficientDimension,-NumDistractors,-Variation) %>%
  mutate(Utterance=gsub("YMax","",as.character(Utterance)))

empirical = tmp_prob %>%
  left_join(tmp_ymin,by=c("SufficientDimension","NumDistractors","Variation","Utterance")) %>%
  left_join(tmp_ymax,by=c("SufficientDimension","NumDistractors","Variation","Utterance")) %>%
  mutate(Utterance = tolower(Utterance),
         Data = "empirical")
```

Load predictives and get into right format.
```{r}
pred = read.csv("../models/bdaOutput/colorSize_predictives.csv",quote="")
colnames(pred)=c("Color","Size","Condition","OtherColor","Item","Utterance","Probability","Zero")
predictives = pred %>%
  mutate(CondLength = nchar(as.character(Condition))) %>%
  mutate(SufficientDimension = 
           case_when( CondLength == 7 ~ "color",
                      CondLength == 6 ~ "size",
                      TRUE ~ "error"),
         NumDistractors = as.numeric(as.character(str_sub(Condition,CondLength-1,CondLength-1))),
         NumSameDistractors = as.numeric(as.character(str_sub(Condition,CondLength,CondLength))),
         NumDiffDistractors = NumDistractors-NumSameDistractors,
         Variation = NumDiffDistractors/NumDistractors,
         Probability = exp(Probability)) %>%
  filter(Utterance %in% c("size","color","size_color")) %>%
  select(SufficientDimension,NumDistractors,Variation,Utterance,Probability) %>%
  mutate(Data = "model",
         YMin = NA,
         YMax = NA)
```

Bind empirical and predictives.
```{r}
d = rbind(empirical,predictives)

redundant = d %>%
  filter(Utterance == "size_color") %>%
  mutate(NumDistractors = as.factor(as.character(NumDistractors)),RedundantDimension=ifelse(SufficientDimension == "color","size redundant","color redundant")) %>%
  droplevels()
```

Make Fig 8 plot analogous to Fig 6.
```{r} 
scale_value=1
ggplot(redundant, aes(x=Variation,y=Probability,color=Data,shape=NumDistractors)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size=.3) +
  geom_point(size=.8) +
  scale_color_manual(values=c(cbPalette[3],cbPalette[5])) +
  xlab("Scene variation") +
  ylab("Probability of redundant modifier") +
  labs(shape="Number of\ndistractors") +
  facet_wrap(~RedundantDimension) +
  theme(panel.border = element_rect(size=.2),
        plot.margin = unit(x = c(0.03, 0.1, 0.03, 0.03), units = "in"),
        panel.grid = element_line(size = .4),
        panel.spacing = unit(2,"pt"),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .7)),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=4,b=4,unit="pt")),
        legend.text      = element_text(margin = margin(l = -2, r=0, unit = "pt"), size = 6 * scale_value),
        legend.title     = element_text(size = 6, face = "bold"),
        legend.position  = "bottom",#c(0.5,-.5),
        legend.direction = "horizontal",
        legend.box = "horizontal",
        legend.spacing.x = unit(.1, 'pt'),
        # legend.box.spacing = unit(2,"pt"),
        legend.margin=margin(0,5,0,0), # change space between sub-legends
        legend.box.margin=margin(-15,0,0,0) # change space between legend and plot
)

# ggsave("../writing/pics/exp1-empirical-predictives.pdf",height=3.9,width=9)
ggsave("../writing/pics/fig8.tiff",units="in",width=2.7,height=1.7,dpi=600)
```

Make Fig 9 a scatterplot of all data points fit in Exp 1.
```{r} 
predictives$ModelProbability = predictives$Probability
predictives$Probability = NULL
scatterd = empirical %>%
  left_join(predictives,by=c("SufficientDimension","NumDistractors","Variation","Utterance")) %>%
  mutate(RedundantDimension=ifelse(SufficientDimension == "color","size redundant","color redundant"))

ggplot(scatterd, aes(x=ModelProbability,y=Probability,color=Utterance,shape=RedundantDimension)) +
  geom_point() +
  scale_color_manual(values=cbPalette) +
  labs(shape="Condition") +
  xlab("MAP model predicted utterance probability") +
  ylab("Empirical utterance proportion") +
  theme(panel.border = element_rect(size=.2),
        # plot.margin = unit(x = c(0.03, 0.1, 0.03, 0.03), units = "in"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .7)),
        legend.text      = element_text(margin = margin(l = -2, r=0, unit = "pt"), size = 6 * scale_value),
        legend.title     = element_text(size = 6, face = "bold"),
        legend.position  = "bottom",#c(0.5,-.5),
        legend.direction = "horizontal",
        legend.box = "vertical",
        legend.spacing.x = unit(1, 'pt'),
        legend.spacing.y = unit(5, 'pt'),
        legend.box.spacing = unit(.1,"pt"),
        legend.margin=margin(0,0,-10,0), # change space between sub-legends
        legend.box.margin=margin(0,0,0,0) # change space between legend and plot
)
# ggsave("../writing/pics/exp1-scatterplot.pdf",height=3,width=4.8)
ggsave("../writing/pics/fig9.tiff",units="in",width=2.7,height=2.8,dpi=600)
```


# Analyze color typicality results

## Parameters

Read parameter data from BDA with **empirical costs** and **empirical semantics** and compute MAPs and HDIs.
```{r}
costs =  "none"#"empirical" # or "empirical"
semantics = "fixedplusempirical" # or "fixed" or "fixedplusempirical"
params <- read_csv(paste("../models/bdaOutput/typicality_cost-",costs,"_sem-",semantics,"_params0.csv",sep="")) %>%
  gather(parameter, value, -logLikelihood, -outputProb)
params %>% 
  group_by(parameter) %>%
  summarise(mode = estimate_mode(value),
            md_lo = round(HPDlo(value), 3),
            md_hi = round(HPDhi(value), 3))
```

Plot posteriors over parameters.
```{r}
paramnames = c("infWeight","fixedVsEmpirical","colorTyp","typeTyp","typWeight")

params.samples = params %>%
  filter(parameter %in% paramnames) %>%
  droplevels() %>%
  mutate(parameter = fct_relevel(parameter,"fixedVsEmpirical","infWeight","typWeight","colorTyp","typeTyp"))

params.samples$parameter2 = factor(params.samples$parameter,
                                   labels = c("beta[fixed]", "beta[i]","beta[t]","x[color]","x[type]"))
                                  
# 
# paramlabeller = labeller(parameter = c(
#   colorTyp = "x[color]",
#   typeTyp = "x[type]",
#   fixedVsEmpirical = "beta[fixed]",
#   typWeight = "beta[t]",
#   infWeight = "beta[i]"))

typParamPosteriors = ggplot(params.samples, aes(x=value))+
    geom_histogram(aes(y=..density..),
               data =subset(params.samples, parameter == "fixedVsEmpirical" ),
               binwidth = .01, colour="black", fill="white") +
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "fixedVsEmpirical"),
                 adjust = 1, alpha=.2, fill="#FF6666")+    
    geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter == "infWeight"),
                   binwidth = .1, colour="black", fill="white") +
    geom_density(aes(y=..density..),
                   data=subset(params.samples, parameter == "infWeight"),
                   adjust = 1, alpha=.2, fill="#FF6666")+
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "typWeight" ),
                 binwidth = .01, colour="black", fill="white") +
    geom_density(aes(y=..density..),
               data=subset(params.samples, parameter == "typWeight"),
               adjust = 1, alpha=.2, fill="#FF6666")+
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "colorTyp" ),
                 binwidth = .005, colour="black", fill="white") +
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "colorTyp"),
                 adjust = 1, alpha=.2, fill="#FF6666")+  
  
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter == "typeTyp" ),
                 binwidth = .001, colour="black", fill="white") +
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "typeTyp"),
                 adjust = 1, alpha=.2, fill="#FF6666")+  

    # ggtitle("Param Posteriors (10,000 samples; lag = 10; burnin=10000)") +
    facet_wrap( ~ parameter2, labeller = label_parsed, scales = 'free',nrow=1) +
  ylab("Density")  + 
  xlab("Value")  + 
  theme(panel.border = element_rect(size=.2),
        plot.margin = unit(x = c(0.03, 0.03, 0.03, 0.03), units = "in"),
        panel.spacing = unit(2, "pt"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .3),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .5)),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=3,b=3,unit="pt"))
)
typParamPosteriors
# ggsave(file=paste("../writing/pics/exp2-cost-",costs,"-sem-",semantics,"-paramposteriors.pdf",sep=""),width=9,height=2)
ggsave("../writing/pics/fig16.tiff",units="in",width=5.7,height=1.3,dpi=600)
```

## Predictives

FIXME Load empirical data and get into right format.

```{r}
empirical = read.csv("../models/bdaInput/typicality/bda_data.csv") %>%
  mutate(target_item = paste(t_color, t_item, sep = "_")) %>%
  mutate(utterance = case_when (
    as.character(response) == as.character(t_color) ~ "color",
    as.character(response) == as.character(t_item) ~ "item",
    as.character(response) == as.character(target_item) ~ "color_item",
    TRUE ~ "other"
  )) %>%
  select(condition,target_item,utterance) %>%
  group_by(condition,target_item) %>%
  count(utterance) %>%
  mutate(EmpiricalProp = n / sum(n)) %>%
  ungroup() %>%
  select(-n)

  # summarize(ModelProb=mean(probability),CILow=ci.low(probability),CIHigh=ci.high(probability)) %>%
  # ungroup() %>%
  # mutate(ModelYMin=ModelProb-CILow,ModelYMax=ModelProb+CIHigh) %>%
  # select(-CILow,-CIHigh)
  
  
  # mutate(color = ifelse(as.character(response) == as.character(t_color),1,0)) %>%
  # mutate(item = ifelse(as.character(response) == as.character(t_item),1,0)) %>%
  # mutate(color_item = ifelse(as.character(response) == as.character(target_item),1,0))

# agr = empirical %>%
#   select(color,item,color_item,condition,target_item) %>%
#   gather(utterance,mentioned,-condition,-target_item) %>%
#   group_by(condition,utterance,target_item) %>%
#   summarise(Probability=mean(mentioned),
#             cilow=ci.low(mentioned),
#             cihigh=ci.high(mentioned)) %>%
#   ungroup() %>%
#   mutate(YMax = Probability + cihigh) %>%
#   mutate(YMin = Probability - cilow) %>%
#   select(condition, target_item, utterance, Probability, YMax, YMin)
# agr = as.data.frame(agr)

```


Load predictives and get into right format.
```{r}
costs = "none"#"empirical" # or "fixed"
semantics = "fixedplusempirical" # or "fixed" or "fixedplusempirical"
# pred = read.csv(paste("../models/bdaOutput/typicality_cost-",costs,"_sem-",semantics,"_predictives.csv",sep=""),quote="")
pred = read.csv("../models/bdaOutput/typicality_none_empirical_predictives0.csv",quote="")
colnames(pred) = c("condition","target_color","target_type","Dist1Color","Dist1Type","Dist2Color","Dist2Type","utterance","logProbability","Zero")
predictives = pred %>%
  mutate(probability = exp(logProbability),
         target_item=paste(target_color,"_",target_type,sep="")) %>%
  # filter(as.character(Utterance) %in% c(as.character(TargetColor),as.character(TargetType),Target)) %>%
  mutate_at(vars(utterance,target_color,target_type),funs(as.character(.))) %>%
  mutate(utterance = case_when(
    utterance == target_color ~ "color",
    utterance == target_type ~ "item",
    utterance == target_item ~ "color_item",
    TRUE ~ "other"
    )) %>%
  filter(utterance != 'other') %>%
  # filter(Utterance == TargetColor | Utterance == TargetType | Utterance == Target) %>%
  select(condition,target_item,utterance,probability) %>%
  group_by(condition,target_item,utterance) %>%
  summarise(ModelProb=mean(probability),CILow=ci.low(probability),CIHigh=ci.high(probability)) %>%
  ungroup() %>%
  mutate(ModelYMin=ModelProb-CILow,ModelYMax=ModelProb+CIHigh) %>%
  select(-CILow,-CIHigh)
```

Load typicalities.
```{r}
typ = read.csv(file="../data/raw/norming_exp2/typicality_exp2_type.csv") %>%
  filter(Utterance == Item) %>%
  mutate(Combo = paste(Color,"_",Item,sep=""))
row.names(typ) = typ$Combo
predictives$type_typicality = typ[as.character(predictives$target_item),]$Typicality
```

Plot model predictions
```{r}

predictives = predictives %>%
  mutate(Utterance = case_when(
    utterance == "item" ~ "type-only",
    utterance == "color" ~ "color-only",
    utterance == "color_item" ~ "color-and-type",
    utterance == "other" ~ "other"))
# change order of levels
predictives$Utterance <- factor(predictives$Utterance, levels=c("type-only", "color-only", "color-and-type", "other"))

modelcaption = expression(paste(bold("(b) "), 
                        "MAP model predicted utterance probabilities"))

modelplot = ggplot(predictives, aes(x=type_typicality,y=ModelProb,color=Utterance)) +
  geom_point(size=.5) +
  geom_smooth(method="lm",se=F) +
  xlab("Typicality of object for type-only utterance") +
  ylab("MAP model predicted probability") +
  scale_color_manual(values = c(cbPalette[2],cbPalette[3],cbPalette[4],"gray30")) +
  facet_wrap(~condition) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  # labs(caption=modelcaption) +
  labs(tag='b') +
  theme(panel.border = element_rect(size=.2),
        # plot.margin = unit(x = c(0.03, 0.1, 0.03, 0.03), units = "in"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .7)),
        legend.text      = element_text(margin = margin(l = -2, r=0, unit = "pt"), size = 6 * scale_value),
        legend.title     = element_text(size = 6, face = "bold"),
        legend.position  = "bottom",#c(0.5,-.5),
        legend.direction = "horizontal",
        legend.box = "vertical",
        legend.spacing.x = unit(1, 'pt'),
        legend.spacing.y = unit(5, 'pt'),
        legend.box.spacing = unit(.1,"pt"),
        legend.margin=margin(0,0,-10,0), # change space between sub-legends
        legend.box.margin=margin(0,0,0,0), # change space between legend and plot
        legend.key=element_blank(),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=3,b=3,unit="pt")),
        plot.tag.position = c(0,1)
        # plot.caption = element_text(size = 7 * scale_value, hjust=.5) #, margin = margin(t = .5)),
)
modelplot
# ggsave(file=paste("../writing/pics/exp2-cost-",costs,"-sem-",semantics,"-paramposteriors.pdf",sep=""),width=9,height=2)
# ggsave("../writing/pics/fig15b.tiff",units="in",width=3.6,height=3,dpi=600)
```

Get empirical plot analogous to model plot, then stick the two together
```{r}
df = read.table(file="../data/data_exp2.csv",sep="\t", header=T, quote="")
agr = df %>%
  select(Color,Type,ColorAndType,Other,NormedTypicality,context) %>%
  gather(Utterance,Mentioned,-context,-NormedTypicality) %>%
  group_by(Utterance,context,NormedTypicality) %>%
  summarise(Probability=mean(Mentioned),ci.low=ci.low(Mentioned),ci.high=ci.high(Mentioned))
agr = as.data.frame(agr)
agr$YMin = agr$Probability - agr$ci.low
agr$YMax = agr$Probability + agr$ci.high
# change order of Utterance column
agr$Utterance <- case_when(
  agr$Utterance == "Type" ~ "type-only",
  agr$Utterance == "Color" ~ "color-only",
  agr$Utterance == "ColorAndType" ~ "color-and-type",
  agr$Utterance == "Other" ~ "other")
# change order of levels
agr$Utterance <- factor(agr$Utterance, levels=c("type-only", "color-only", "color-and-type", "other"))
# change context names to have nicer facet labels 
levels(agr$context) = c("informative","informative-cc", "overinformative", "overinformative-cc")

# make banana item data.frame to circle
bananas = df %>%
  filter(Item == "banana") %>%
  select(Color,Type,ColorAndType,Other,NormedTypicality,context,TargetColor) %>%
  gather(Utterance,Mentioned,-context,-NormedTypicality,-TargetColor) %>%
  group_by(Utterance,context,NormedTypicality,TargetColor) %>%
  summarise(Probability=mean(Mentioned)) %>%
  ungroup() %>%
  filter(Utterance == "ColorAndType") %>%
  mutate(X=NormedTypicality,Y=Probability,Utterance="color-and-type") %>%
  select(context,TargetColor,X,Y,Utterance)

# plot
scale_value = 1
empiricalcaption = expression(paste(bold("(a) "), 
                        "Empirical utterance proportions"))
empiricalplot = ggplot(agr, aes(x=NormedTypicality,y=Probability,color=Utterance)) +
  geom_point(size=.5) +
  geom_smooth(method="lm",se=F) +
  scale_color_manual(values = c(cbPalette[2],cbPalette[3],cbPalette[4],"gray30","red","orange")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  new_scale_color() +
  geom_point(data=bananas,aes(x=X,y=Y,color=TargetColor),shape=0,size=3,stroke=1,show.legend=F) +
  scale_color_manual(values = c("dodgerblue3","chocolate4","gold")) +
  facet_wrap(~context) +
  xlab("Typicality of object for type-only utterance") +
  ylab("Empirical utterance proportion") +
  # labs(caption=empiricalcaption) +
  labs(tag = 'a') +
  coord_cartesian(xlim=c(0.4,1),ylim=c(0, 1)) +
  theme(panel.border = element_rect(size=.2),
        # plot.margin = unit(x = c(0.03, 0.1, 0.03, 0.03), units = "in"),
        panel.grid = element_line(size = .4),
        axis.line        = element_line(colour = "black", size = .2),
        axis.ticks       = element_line(colour = "black", size = .2),
        axis.ticks.length = unit(2, "pt"),
        axis.text.x        = element_text(size = 6 * scale_value, colour = "black",vjust=2),
        axis.text.y        = element_text(size = 6 * scale_value, colour = "black",margin = margin(r = 0.3)),#,hjust=-5),
        axis.title.x       = element_text(size = 6 * scale_value, margin = margin(t = .5)),
        axis.title.y       = element_text(size = 6 * scale_value, margin = margin(r = .7)),
        legend.text      = element_text(margin = margin(l = -2, r=0, unit = "pt"), size = 6 * scale_value),
        legend.title     = element_text(size = 6, face = "bold"),
        legend.position  = "bottom",#c(0.5,-.5),
        legend.direction = "horizontal",
        legend.box = "vertical",
        legend.spacing.x = unit(1, 'pt'),
        legend.spacing.y = unit(5, 'pt'),
        legend.box.spacing = unit(.1,"pt"),
        legend.margin=margin(0,0,-10,0), # change space between sub-legends
        legend.box.margin=margin(0,0,0,0), # change space between legend and plot
        # legend.key=element_blank(),
        # legend.key = element_rect(colour = "transparent", fill = "transparent"),
        strip.text      = element_text(size = 6 * scale_value,margin=margin(t=3,b=3,unit="pt")),
        plot.tag.position = c(0,1)
        # plot.caption = element_text(size = 7 * scale_value, hjust=.5) #, margin = margin(t = .5)),
  )
empiricalplot

tiff(file="../writing/pics/fig15.tiff",units="in",width=3.6,height=6.5,res=600)
grid.arrange(empiricalplot,modelplot,nrow=2)
dev.off()

```

Bind empirical and predictives and plot correlation
```{r}
d = left_join(predictives,empirical,by=c("condition","target_item","utterance")) %>%
  replace_na(list(EmpiricalProp = 0))
cor(d$ModelProb,d$EmpiricalProp)

ggplot(d, aes(x=ModelProb,y=EmpiricalProp,color=condition)) +
  geom_errorbar(aes(ymin=ModelYMin,ymax=ModelYMax)) +
  # geom_errorbarh(aes(xmin=ModelYMin,xmax=ModelYMax)) +
  geom_point() +
  xlab("Model predicted probability") +
  ylab("Empirical utterance proportion") +
  # facet_wrap(~condition) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  guides(color=guide_legend(title="utterance")) +
  theme(aspect.ratio = 1)
ggsave("../writing/pics/exp3-empirical-predictives.pdf",height=4,width=11)
```

## Model comparison

Load BF results

```{r}
AIS.samples <- read_csv('../models/BFOutput/BF_typicality.txt') %>%
  group_by(modelVersion, costs, semantics) %>%
  summarize(m = mean(margLikelihood))

# fixedplusempirical vs. empirical
AIS.samples %>% spread(semantics, m) %>%
  mutate(BF1 = exp(fixedplusempirical - empirical),
         BF2 = format( exp(fixedplusempirical - fixed),
                       scientific = TRUE, digits = 3))

# no cost vs. cost
AIS.samples %>% filter(semantics == 'fixedplusempirical') %>% 
  spread(costs, m) %>%
  mutate(noneVsFixedBF = exp(none - fixed)) %>%
  mutate(noneVsEmpBF = exp(none - empirical)) %>%
  select(noneVsFixedBF, noneVsEmpBF)
```

# Analyze nominal choice results

## Parameters

Read parameter data from BDA and compute MAPs and HDIs for cost-model.
```{r}
params <- read_csv("../models/bdaOutput/nominal_cost-empirical_sem-truthconditionalplusempirical_params1.csv") %>%
  gather(parameter, value, infWeight:empiricalVsTruthConditional)

params %>% 
  group_by(parameter) %>%
  summarize(mode = estimate_mode(value),
            md_lo = round(HPDlo(value), 3),
            md_hi = round(HPDhi(value), 3))
```

```{r}
params.samples <- getSamples(params)
params.samples$parameter2 = factor(params.samples$parameter,
                                   labels = c("beta[fixed]","beta[i]","beta[t]","beta[F]","beta[L]"),
                                   levels = c("empiricalVsTruthConditional","infWeight","typWeight","freqCostWeight","lengthCostWeight")) 

params.samples %<>%
  # filter(parameter != "empiricalVsTruthConditional") %>%
  droplevels()
  
nomParamPosteriors = ggplot(params.samples, aes(x=value, alpha = .2))+
    geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter2 == "beta[fixed]"),
                   binwidth = .001, colour="black", fill="white") +
      geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter2 == "beta[i]"),
                   binwidth = .1, colour="black", fill="white") +
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter2 == "beta[t]" ),
                 binwidth = .01, colour="black", fill="white") +  
    geom_histogram(aes(y=..density..),
                 data =subset(params.samples, parameter2 == "beta[L]" ),
                 binwidth = .05, colour="black", fill="white") +
      geom_histogram(aes(y=..density..),
                   data=subset(params.samples, parameter2 == "beta[F]"),
                   binwidth = .01, colour="black", fill="white") +
    #geom_histogram(aes(y=..density..),
     #            data =subset(params.samples, parameter2 == "semanticInterp" ),
      #           binwidth = .1, colour="black", fill="white") +

    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter2 == "beta[fixed]"),
                 adjust = 1, alpha=.2, fill="#FF6666") +
      geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter2 == "beta[i]"),
                 adjust = 1, alpha=.2, fill="#FF6666") +
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter2 == "beta[t]"),
                 adjust = 1, alpha=.2, fill="#FF6666") +  
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter2 == "beta[L]"),
                 adjust = 1, alpha=.2, fill="#FF6666") +
      geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter2 == "beta[F]"),
                 adjust = 1, alpha=.2, fill="#FF6666") +
   #   geom_density(aes(y=..density..),
    #             data=subset(params.samples, parameter2 == "semanticInterp"),
     #            adjust = 1, alpha=.2, fill="#FF6666") +

    # ggtitle("Param Posteriors (2000 samples; lag = 10; burnin=10000)") +
    facet_wrap( ~ parameter2, scales = 'free',nrow=1,labeller = label_parsed) +
    theme_bw() +
  theme(legend.position="none") +
    ylab("Density") 
nomParamPosteriors
ggsave(file="../writing/pics/exp3-paramposteriors.pdf",width=9,height=2)
```



## Predictives

Load empirical data and get into right format.

```{r}
emp = read.csv("../models/bdaInput/nominal/bda_data.csv")
colnames(emp) = c("Pair","Trial","Cond","Target","Dist1","Dist2","Utterance")

empirical = emp %>%
  mutate(Condition = 
           case_when(Cond == "basic12" ~ "sub necessary",
                     Cond == "basic33" ~ "super sufficient",
                     TRUE ~ "basic sufficient"),
         sub = Utterance == "sub",
         basic = Utterance == "basic",
         super = Utterance == "super") %>%
  group_by(Condition, Target) %>%
  summarize(ProbSub = mean(sub),ProbBasic = mean(basic),ProbSuper = mean(super),CILowSub=ci.low(sub),CIHighSub=ci.high(sub),CILowBasic=ci.low(basic),CIHighBasic=ci.high(basic),CILowSuper=ci.low(super),CIHighSuper=ci.high(super)) %>%
  ungroup() %>%
  mutate(YMinSub=ProbSub-CILowSub,YMaxSub=ProbSub+CIHighSub,YMinBasic=ProbBasic-CILowBasic,YMaxBasic=ProbBasic+CIHighBasic,YMinSuper=ProbSuper-CILowSuper,YMaxSuper=ProbSuper+CIHighSuper) 

tmp_prob = empirical %>%
  select(Condition,Target,ProbSub,ProbBasic,ProbSuper) %>%
  gather(Utterance, Probability, -Condition, -Target) %>%
  mutate(Utterance=gsub("Prob","",as.character(Utterance)))

tmp_ymin = empirical %>%
  select(Condition,Target,YMinSub,YMinBasic,YMinSuper) %>%
  gather(Utterance, YMin,-Condition,-Target) %>%
  mutate(Utterance=gsub("YMin","",as.character(Utterance)))

tmp_ymax = empirical %>%
  select(Condition,Target,YMaxSub,YMaxBasic,YMaxSuper) %>%
  gather(Utterance, YMax,-Condition,-Target) %>%
  mutate(Utterance=gsub("YMax","",as.character(Utterance)))

empirical = tmp_prob %>%
  left_join(tmp_ymin,by=c("Condition","Target","Utterance")) %>%
  left_join(tmp_ymax,by=c("Condition","Target","Utterance")) %>%
  mutate(Utterance = tolower(Utterance))
```

Load predictives and get into right format.
```{r}
pred = read.csv("../models/bdaOutput/nominal_cost-empirical_sem-truthconditionalplusempirical_predictives1.csv",quote="")
colnames(pred)=c("Cond","Target","Dist1","Dist2","Utt","logProbability","Zero") 
predictives = pred %>%
    mutate(Condition = 
           case_when(Cond == "basic12" ~ "sub necessary",
                     Cond == "basic33" ~ "super sufficient",
                     TRUE ~ "basic sufficient"),
           Probability = exp(logProbability),
           Utterance = 
             case_when(Utt %in% c("vehicle","clothing","furniture","animal","snack","plant") ~ "super",
                       Utt %in% c("bear","bird","candy","car","dog","fish","flower","shirt","table") ~ "basic",
                       TRUE ~ "sub")) %>%
  select(Condition,Target,Utterance,Probability) %>%
  group_by(Condition,Target,Utterance) %>%
  summarize(ModelProb=mean(Probability),CILow=ci.low(Probability),CIHigh=ci.high(Probability)) %>%
  ungroup() %>%
  mutate(ModelYMin=ModelProb-CILow,ModelYMax=ModelProb+CIHigh) %>%
  select(-CILow,-CIHigh)
```

Bind empirical and predictives and create scatterplot.
```{r}
d = left_join(empirical,predictives,by=c("Condition","Target","Utterance")) %>%
  mutate(Condition = relevel(as.factor(Condition), "sub necessary"))

# correlation at the condition-target-utterance level (324 data points)
cor(d$ModelProb,d$Probability)

# correlation at the condition-utterance level
cors = d %>%
  group_by(Condition,Utterance) %>%
  summarize(MeanModel=mean(ModelProb),MeanEmp=mean(Probability))
cor(cors$MeanModel,cors$MeanEmp)

ggplot(d, aes(x=ModelProb,y=Probability,color=relevel(as.factor(Utterance),"sub"))) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax)) +
  geom_errorbarh(aes(xmin=ModelYMin,xmax=ModelYMax)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  xlab("Model predicted probability") +
  ylab("Empirical utterance proportion") +
  facet_wrap(~Condition) +
  # facet_wrap(~relevel(as.factor(Condition),"sub necessary")) +
  guides(color=guide_legend(title="Utterance")) +
  theme_bw() +
  theme(aspect.ratio = 1) 
ggsave("../writing/pics/exp3-empirical-predictives.pdf",height=2.5,width=7)
```


Plot model predictions and empirical means as a function of typicality. First add typicality values, then make same plot as Fig 19 but with model predictions instead of empirical
```{r}
typs = read.table("../data/typicality_exp3.csv",header=T,quote="",sep=",")
ttyps = droplevels(subset(typs, itemtype == "target"))
row.names(ttyps) = paste(ttyps$labeltype, ttyps$item)
d$typ_sub = ttyps[paste("sub",as.character(d$Target)),]$meanresponse
d$typ_basic = ttyps[paste("basic",as.character(d$Target)),]$meanresponse
d$typ_super = ttyps[paste("super",as.character(d$Target)),]$meanresponse

toplot = d %>%
  select(Condition, Target, Utterance, typ_sub, Probability, ModelProb) %>%
  gather(Data, Probability, Probability:ModelProb) %>%
  mutate(Data = fct_recode(Data,empirical="Probability",model="ModelProb"))

toplot$item = ""
toplot[toplot$Target == "parrot" & toplot$Data == "empirical",]$item = "parrot"
toplot[toplot$Target == "poloshirt" & toplot$Data == "empirical",]$item = "poloshirt"
toplot[toplot$Target == "suv" & toplot$Data == "empirical",]$item = "suv"

ggplot(toplot, aes(x=typ_sub,y=Probability,color=relevel(as.factor(Utterance),"sub"))) +
  geom_point() +
  geom_smooth(method="lm") +
  scale_color_manual(values=cbPalette[2:4]) +
  geom_text(aes(label=item),nudge_y=.08) +
  ylab("Utterance probability") +
  xlab("Typicality of object for subordinate level utterance") +
  facet_grid(relevel(Data,"empirical")~relevel(as.factor(Condition),"sub necessary")) +
  guides(color=guide_legend(title="Utterance"))
ggsave("../writing/pics/results-exp3.png",height=4,width=7.5)
```

## Model comparison

Load BF results

```{r}
AIS.samples <- read_csv('../models/BFOutput/BF_nominal.txt', col_names = c('modelVersion', 'costs', 'semantics', 'margLikelihood')) %>%
  group_by(modelVersion, costs, semantics) %>%
  summarize(m = mean(margLikelihood))

# no cost vs. cost
AIS.samples %>% 
  spread(costs, m) %>%
  mutate(noneVsEmpBF = exp(empirical - none)) %>%
  select(noneVsEmpBF)
```

